---
title: "How does first programming language affect ability to learn R?"
output: github_document
---

**Author:** John Blischak

**Date:** 2020-04-22

**License:** [CC-BY 4.0](https://creativecommons.org/licenses/by/4.0/)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, fig.width = 10)
```

```{r packages}
library(tidyverse)
theme_set(theme_classic(base_size = 18))
```

```{r import}
survey_url <- "https://raw.githubusercontent.com/rstudio/learning-r-survey/master/2018/data/survey_English.tsv"
survey <- read_delim(survey_url, delim = "\t")
```

The data comes from [RStudio's 2018 survey on learning
R](https://github.com/rstudio/learning-r-survey). The raw data includes
responses from `r nrow(survey)` participants.

```{r first-language, results='hide'}
sum(is.na(survey$Qfirst_language))
survey <- drop_na(survey, Qfirst_language)
r <- grep("^R$", survey$Qfirst_language, ignore.case = TRUE, value = TRUE)
first_r <- survey$Qfirst_language %in% r
sum(first_r)

gui <- c(
  grep("Excel", survey$Qfirst_language, ignore.case = TRUE, value = TRUE),
  grep("Minitab", survey$Qfirst_language, ignore.case = TRUE, value = TRUE),
  grep("SPSS", survey$Qfirst_language, ignore.case = TRUE, value = TRUE),
  grep("Tableau", survey$Qfirst_language, ignore.case = TRUE, value = TRUE)
)
first_gui <- survey$Qfirst_language %in% gui
sum(first_gui)

basic <- grep("basic", survey$Qfirst_language, ignore.case = TRUE, value = TRUE)
sort(table(basic), decreasing = TRUE)[1:5]
first_basic <- survey$Qfirst_language %in% basic
sum(first_basic)

data_centric <- c(
  grep("maple", survey$Qfirst_language, ignore.case = TRUE, value = TRUE),
  grep("mathematica", survey$Qfirst_language, ignore.case = TRUE, value = TRUE),
  grep("matlab", survey$Qfirst_language, ignore.case = TRUE, value = TRUE),
  grep("SAS", survey$Qfirst_language, ignore.case = TRUE, value = TRUE),
  grep("Stata", survey$Qfirst_language, ignore.case = TRUE, value = TRUE),
  grep("Lisrel", survey$Qfirst_language, ignore.case = TRUE, value = TRUE)
)
sort(table(data_centric), decreasing = TRUE)[1:5]
first_data_centric <- survey$Qfirst_language %in% data_centric
sum(first_data_centric)

scripting <- c(
  grep("perl", survey$Qfirst_language, ignore.case = TRUE, value = TRUE),
  grep("php", survey$Qfirst_language, ignore.case = TRUE, value = TRUE),
  grep("python", survey$Qfirst_language, ignore.case = TRUE, value = TRUE),
  grep("ruby", survey$Qfirst_language, ignore.case = TRUE, value = TRUE)
)
sort(table(scripting), decreasing = TRUE)[1:4]
first_scripting <- survey$Qfirst_language %in% scripting
sum(first_scripting)

lisp <- c(
  grep("lisp", survey$Qfirst_language, ignore.case = TRUE, value = TRUE),
  grep("scheme", survey$Qfirst_language, ignore.case = TRUE, value = TRUE),
  grep("racket", survey$Qfirst_language, ignore.case = TRUE, value = TRUE)
)
sort(table(lisp), decreasing = TRUE)[1:5]
first_lisp <- survey$Qfirst_language %in% lisp
sum(first_lisp)

c_plus_plus <- c(
  grep("C\\+\\+", survey$Qfirst_language, ignore.case = TRUE, value = TRUE),
  grep("^C$", survey$Qfirst_language, ignore.case = TRUE, value = TRUE)
)
sort(table(c_plus_plus), decreasing = TRUE)
first_c_plus_plus <- survey$Qfirst_language %in% c_plus_plus
sum(first_c_plus_plus)

java <- c(
  grep("^java$", survey$Qfirst_language, ignore.case = TRUE, value = TRUE)
)
sort(table(java), decreasing = TRUE)
first_java <- survey$Qfirst_language %in% java
sum(first_java)

fortran <- c(
  grep("fortran", survey$Qfirst_language, ignore.case = TRUE, value = TRUE)
)
sort(table(fortran), decreasing = TRUE)[1:5]
first_fortran <- survey$Qfirst_language %in% fortran
sum(first_fortran)

pascal <- c(
  grep("pascal", survey$Qfirst_language, ignore.case = TRUE, value = TRUE)
)
sort(table(pascal), decreasing = TRUE)[1:5]
first_pascal <- survey$Qfirst_language %in% pascal
sum(first_pascal)

survey$first_lang <- dplyr::case_when(
  first_basic & !first_c_plus_plus & !first_data_centric & !first_fortran &
    !first_gui & !first_java & !first_lisp & !first_pascal & !first_r & !first_scripting ~ "BASIC",
  !first_basic & first_c_plus_plus & !first_data_centric & !first_fortran &
    !first_gui & !first_java & !first_lisp & !first_pascal & !first_r & !first_scripting ~ "C/C++",
  !first_basic & !first_c_plus_plus & first_data_centric & !first_fortran &
    !first_gui & !first_java & !first_lisp & !first_pascal & !first_r & !first_scripting ~ "Data-centric",
  !first_basic & !first_c_plus_plus & !first_data_centric & first_fortran &
    !first_gui & !first_java & !first_lisp & !first_pascal & !first_r & !first_scripting ~ "Fortran",
  !first_basic & !first_c_plus_plus & !first_data_centric & !first_fortran &
    first_gui & !first_java & !first_lisp & !first_pascal & !first_r & !first_scripting ~ "GUI",
  !first_basic & !first_c_plus_plus & !first_data_centric & !first_fortran &
    !first_gui & first_java & !first_lisp & !first_pascal & !first_r & !first_scripting ~ "Java",
  !first_basic & !first_c_plus_plus & !first_data_centric & !first_fortran &
    !first_gui & !first_java & first_lisp & !first_pascal & !first_r & !first_scripting ~ "Lisp",
  !first_basic & !first_c_plus_plus & !first_data_centric & !first_fortran &
    !first_gui & !first_java & !first_lisp & first_pascal & !first_r & !first_scripting ~ "Pascal",
  !first_basic & !first_c_plus_plus & !first_data_centric & !first_fortran &
    !first_gui & !first_java & !first_lisp & !first_pascal & first_r & !first_scripting ~ "R",
  !first_basic & !first_c_plus_plus & !first_data_centric & !first_fortran &
    !first_gui & !first_java & !first_lisp & !first_pascal & !first_r & first_scripting ~ "Scripting"
)
table(survey$first_lang, useNA = "ifany")
survey <- drop_na(survey, first_lang)
nrow(survey)
```

```{r wrangling}
survey <- survey %>%
  drop_na(
    Qr_year,
    Qyear_born,
    Qr_experience,
    Qtime_to_proficiency,
    Qr_enjoyment,
    Qr_difficulty_experienced,
    Qrecommend
  ) %>%
  filter(
    # S was released in 1976
    Qr_year >= 1976, 
    Qr_year <= 2018,
    Qyear_born >= 1900,
    Qyear_born <= 2000,
    # Have to have some experience with R
    Qr_experience != "None"
  ) %>%
  mutate(
    age = 2018 - Qyear_born,
    r_years = 2018 - Qr_year,
    Qr_experience = factor(Qr_experience,
                           levels = c("Beginner", "Intermediate", "Expert")),
    Qtime_to_proficiency = factor(Qtime_to_proficiency,
                                  levels = c("I don't feel proficient in R yet.",
                                             "Days", "Weeks", "Months", "Years"))
  )
```

I focused on the data of [reported first language
learned](https://github.com/rstudio/learning-r-survey/blob/master/2018/dictionary/first-language-dictionary.tsv).
Specifically, I was interested to see if there was any evidence that prior
exposure to specific programming paradigms influenced the ability to learn R.
For example, since [R began as a Lisp-like interpreter written in
C](https://www.stat.auckland.ac.nz/~ihaka/downloads/Interface98.pdf), does
experience with Lisp and/or C provide any advantage when trying to grok R? After
applying quality filters and only keeping participants that could be assigned to
a large enough group of first programming language, I was left with
`r nrow(survey)` participants.

**Disclaimers:**

* I'm a geneticist. Don't take any of this too seriously.

* The study authors themselves warn of the incredible sampling bias. And as you
will see from the results below, the participants are satisfied R users.

## First programming language

```{r first-lang}
survey %>%
  ggplot() +
  aes(x = first_lang) +
  geom_bar() +
  labs(x = "First programming language",
       y = "Number of participants")
```

## Demographics

But isn't first programming language going to be super confounded with age and
experience? Surprisingly, not as much as I had expected.

While Fortran-first programmers tend to be older, the remaining categories are
comparable.

```{r age}
survey %>%
  ggplot() +
  aes(x = first_lang, y = age) +
  geom_boxplot() +
  labs(x = "First programming language",
       y = "Age")
```

Looking at years of R usage, Fortran-first users had more experience as
expected. But Lisp-first programmers had the highest median years of R usage.

```{r r-usage}
survey %>%
  ggplot() +
  aes(x = first_lang, y = r_years) +
  geom_boxplot() +
  labs(x = "First programming language",
       y = "Years of R usage")
```

And for self-reported R experience, the main outlier are those whose first
programming environment was a GUI (e.g. Excel, SPSS), with a larger percentage
of self-reported beginners.

```{r r-experience}
survey %>%
  group_by(first_lang) %>%
  count(Qr_experience) %>%
  mutate(perc = n / sum(n)) %>%
  ggplot() +
  aes(x = first_lang, y = perc, fill = Qr_experience) +
  geom_col() +
  scale_fill_brewer("R experience", palette = "Dark2") +
  theme(legend.position = "top") +
  labs(x = "First programming language",
       y = "Percentage")
```

## Roughly how long did it take you to achieve proficiency in R?

No super clear advantage for time to proficiency.

```{r time-to-proficiency}
survey %>%
  group_by(first_lang) %>%
  count(Qtime_to_proficiency) %>%
  mutate(perc = n / sum(n)) %>%
ggplot() +
  aes(x = first_lang, y = perc, fill = Qtime_to_proficiency) +
  geom_col() +
  scale_fill_brewer("Time to proficiency", palette = "Dark2") +
  theme(legend.position = "top") +
  labs(x = "First programming language",
       y = "Percentage")
```

## Please rate how much you enjoy using R on a scale of 1 to 5, where 1 is you don't enjoy it at all, and 5 is that you enjoy it a great deal.

Everybody loves R.

```{r enjoy}
survey %>%
  ggplot() +
  aes(x = first_lang, y = Qr_enjoyment) +
  geom_boxplot() +
  labs(x = "First programming language",
       y = "Enjoyment (1 - 5)")
```

## Compared with other technical topics you've learned in school and on the job, on a scale of 1 to 5, how difficult has it been for you to learn R?

GUI-first and R-first participants find it most difficult to learn R. Lisp-first
had the easiest time (though note they were the smallest sample). So really
learning any programming language appears to help with learning R.

```{r difficulty}
survey %>%
  ggplot() +
  aes(x = first_lang, y = Qr_difficulty_experienced) +
  geom_boxplot() +
  labs(x = "First programming language",
       y = "Difficulty (1 - 5)")
```

## How likely are you to recommend R to a colleague, friend, or family member?

Most participants highly recommend R. Especially the Lisp-first participants.
And the Scripting-first participants had the largest number of lower likeliness
to recommend R, but this group includes Python, so maybe that is to be expected?
:-)

```{r recommend}
survey %>%
  ggplot() +
  aes(x = first_lang, y = Qrecommend) +
  geom_boxplot() +
  labs(x = "First programming language",
       y = "Difficulty (1 - 10)")
```

## Session information

```{r session-info}
sessionInfo()
```
